---
title: "Reporting Analyst Evaluation"
author: "Cici Chen"
date: "10/24/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- You may need to install these packages if you do not have them in your R. -->
```{r install.packages, echo=F, warning=F, message=F}
if(!require("readxl")) install.packages("readxl")
if(!require("ggplot2")) install.packages("ggplot2")
if(!require("lubridate")) install.packages("lubridate")
if(!require("plotrix")) install.packages("plotrix")
if(!require("packcircles")) install.packages("packcircles")
if(!require("webshot")) install.packages("webshot")
if(!require("kableExtra")) install.packages("kableExtra")
webshot::install_phantomjs()
```

```{r libraries, echo=F,message=F, warning=F}
library(DT)
library(readxl)
library(datamodelr)
library(readxl)
library(ggplot2)
library(lubridate)
library(plotrix)
library(plyr)
library(knitr)
library(kableExtra)
library(scales)
library(grid)
library(dplyr)
library(packcircles)
```

```{r functions, echo=F}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Part 1: Entity Relationship Diagram

ERD could be shown as below:

```{r connect,echo=F}
sheetNum <- excel_sheets("Adventure Works Data.xlsx")

# Read in 6 sheets
Employee.HR<-read_excel("Adventure Works Data.xlsx",sheet=sheetNum[1])
Business.Entity.Address<-read_excel("Adventure Works Data.xlsx",sheet=sheetNum[2])
Sales.person<-read_excel("Adventure Works Data.xlsx",sheet=sheetNum[3])
Contact<-read_excel("Adventure Works Data.xlsx",sheet=sheetNum[4])
Pay.history<-read_excel("Adventure Works Data.xlsx",sheet=sheetNum[5])
Address<-read_excel("Adventure Works Data.xlsx",sheet=sheetNum[6])
```

```{r ERD, echo=F}
dm_f <- dm_from_data_frames(Employee.HR,Business.Entity.Address,Sales.person,Contact,Pay.history,Address)%>%
        dm_add_references(
           Pay.history$`Business Entity ID`==Employee.HR$`Business Entity ID`,
           Sales.person$`Business Entity ID`==Employee.HR$`Business Entity ID`,
           Business.Entity.Address$`Business Entity ID`==Address$`Address ID`,
           Employee.HR$`Business Entity ID`==Contact$`BusinessEntityID (Person)`,
           Business.Entity.Address$`Address ID`==Employee.HR$`Business Entity ID`)%>%
  dm_set_key("Business.Entity.Address", c("Address ID","Business Entity ID"))%>%
  dm_set_key("Pay.history", "Business Entity ID")%>%
  dm_set_key("Sales.person", "Business Entity ID")

graph <- dm_create_graph(dm_f, rankdir = "BT",col_attr = c("column", "type"))
dm_render_graph(graph)
```

## Part 2: Connecting to Data
Now connect the Adventure Works data file. As I also want to include those employees who are not sales persons, I use **Left Join** for different sheets.

```{r merge, echo=F}
# Merge the sheets we need
a<-merge(Employee.HR,Sales.person,by="Business Entity ID",all.x =TRUE) # left join 
b<-merge(a,Contact,by.x = "Business Entity ID",by.y="BusinessEntityID (Person)",all.x=T) # left join 
c<-merge(b,Pay.history,by="Business Entity ID",all.x=T) # left join 
```

## Part 3: Data Tables (Cross Tabs)
I created a list of the employees with their *Business Id, Last Name, First Name, Job Title, Rate, Bonus, Sales Quota and Sales Last Year*. Format this sheet to include the following:
```{r create.a.list,echo=F}
dat<-data.frame(c$`Business Entity ID`,c$`Last Name`,c$`First Name`,c$`Job Title`,c$Gender,
                c$`Hire Date`,c$Rate,c$Bonus,c$`Sales Quota`,c$`Sales Last Year`,c$`Sales YTD`)
colnames(dat)<-c("Business Id","Last Name","First Name","Job Title","Gender","Hire Date",
                   "Rate", "Bonus","Sales Quota","Sales Last Year","Sales Year to Date")
```

##### 1. Sort by Salary Rate
```{r sort.by.salary, echo=F}
dat<-dat[order(dat$Rate,decreasing = T),]

kable(head(dat,15), "latex", booktabs = T, row.names = F) %>% 
  kable_styling(latex_options = "scale_down",font_size = 15)
```

##### 2. Total the Columns:
```{r totals, echo=F,eval=F}
Totals<-apply(dat[c("Rate" ,"Bonus","Sales Quota","Sales Last Year" )],2,sum, na.rm=T)
t(as.data.frame(Totals))
```

||Rate |Bonus| Sales Quota |Sales Last Year|
|-|-|-|-|-|
|Totals| 5274.594| 48610   |  3650000   |  23685964|

##### 3. What else can you do to make this table more readable? Please implement your ideas

* Change Gender from **"F & M"** to **"Female & Male"**
* **Round** the Rate
* The last 4 columns *Bonus, Sales Quota, Sales Last Year* are actually **ONLY** for **Sales Department**, so we need to make a comment under the table for better interpretation

Then we can have a more readable table:
```{r fancy.table, echo=F}
dat$Gender<-ifelse(dat$Gender=="F","Female","Male")
dat$Gender<-as.factor(dat$Gender)
dat$Rate<-round(dat$Rate,1)

kable(head(dat,15), "latex", booktabs = T, row.names = F) %>% 
  kable_styling(latex_options = c("striped", "scale_down"),font_size = 15)
```

**Comments**: 

* *Bonus, Sales Quota, Sales Last Year* are only for *Sales Department*

* **"NA"** stands for **"Unknown / Not Applicable / Missing Values"**

##### 4. Questions: 

* Q1. **Ken Sánchez**	has the higest salary rate in the firm and she is **Chief Executive Officer**.
```{r higest.rate, echo=F, eval=F}
dat[order(dat$Rate,decreasing = T),][1,c(1,2,3,4,7)]
```

* Q2. **Tsvi Reiter**	has the higest salary Bonus last year and it is **\$6,700**.
```{r higest.bonus,echo=F, eval=F}
dat[order(dat$Bonus, decreasing = T),][1,c(1,2,3,8)]
```

* Q3. **Ranjit Varkey Chudukatil**	has the higest sales last year and it is **\$2,396,540**.
```{r higest.sales.last.year, echo=F, eval=F}
dat[order(dat$`Sales Last Year`,decreasing = T),][1,c(1,2,3,10)]
```

* Q4. **Tete Mensa-Annan Tete** does not meet her quota.
```{r under.quota, echo=F,eval=F}
dat[which(dat$`Sales Quota`>dat$`Sales Last Year`),][,c(1,2,3)]
```

* Q5. 

|Summary|Rate|
|--|--|
|Mean|18.18|
|Median|14|
|Mode|9.5|

```{r summary.mean, echo=F}
mean<-paste("Mean =",round(mean(dat$Rate),2))
median<-paste("Median =",median(dat$Rate))
#table(dat$Rate)
mode<-paste("Mode =",names(table(dat$Rate))[which(table(dat$Rate)==max(table(dat$Rate)))])
```

##### 5: Q6. What other interesting insights can be derived?
##### Insight 1: Gender vs. Average Rate ==> More Male while Female tends to earn more
```{r Insight1.1,echo=F}
gender.dim<-ddply(dat, "Gender", summarise, Mean=round(mean(Rate),2),Median=median(Rate))
percentage<-100*round(table(dat$Gender)/(206+84),4)
```

```{r Insight1.2,fig.align='center',fig.width=7,fig.height=4, echo=F}
# Pie
data <- data.frame(
  group=c("28.97% Female","71.03% Male"),
  value=c(table(dat$Gender)[[1]],table(dat$Gender)[[2]])
)

data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = value / sum(data$value) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )    # Compute the position of labels

p1<-ggplot(data, aes(x="", y=prop, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +labs(title="Gender vs. Average Rate") +
  coord_polar("y", start=0) +
  theme_void() + 
  theme(legend.position="bottom",legend.title=element_blank(),plot.title = element_text(hjust = 0.8)) +
  geom_text(aes(y = ypos-3, label = c("Average Rate=$17.76","Average Rate=$19.22")), color = "white", size=3) +
  scale_fill_manual(values=c("hotpink1","steelblue2"))

# Density
cdat<-ddply(dat, "Gender", summarise, rating.mean=mean(Rate))
p2<-ggplot(dat, aes(x=Rate, colour=Gender,fill=Gender)) +
    geom_density(alpha=.3) +
    geom_vline(data=cdat, aes(xintercept=rating.mean, colour=Gender),linetype="dashed", size=1) +
    theme(legend.position="top",legend.title=element_blank()) +
    labs(x="Rate",y="")+
  scale_fill_manual(values=c("hotpink1", "steelblue2"))
  
multiplot(p1, p2, cols=2)
```

* The employees are composed of **28.97%** Females and **71.03%** Males
* The average rate of Females(**\$19.22**) is larger than the average rate of Males(**\$17.76**)
* Females seem to earn more than Males, while the largest rate(**\$125**) is from a Male

##### Insight 2: Peak Period vs. Hire Date ==> Time of a year influences the hiring of the firm
```{r insight2, fig.align='center',fig.width=7,fig.height=3,echo=F}
yrs<-format(as.Date(dat$`Hire Date`), "%Y")
yrs.freq<-data.frame(table(yrs))
ms<-format(as.Date(dat$`Hire Date`), "%m")
ms.freq<-data.frame(table(ms))

yrs.freq$yrs<-as.numeric(as.character(yrs.freq$yrs))
ms.freq$ms<-as.numeric(as.character(ms.freq$ms))

p3<-ggplot(yrs.freq, aes(x=yrs, y=Freq)) +
  geom_line( color="#69b3a2", size=1, alpha=0.9, linetype=1) +
  labs(title="Peak Period of Hirng season",x="Years",y="Count of hiring numbers")+
  theme(plot.title = element_text(hjust = 0.5))

p4<-ggplot(ms.freq, aes(x=ms, y=Freq)) +
  geom_line( color="#69b3a2", size=1, alpha=0.9, linetype=1) +
  labs(title="Seasonal Change of Peak Period",x="Months",y="")+
  theme(plot.title = element_text(hjust = 0.5))

multiplot(p3, p4, cols=2)
```

* From the left graph, the firm hired a lot of people (around **200**) in **Year=2003** , and there is a peak for the hiring process based on different years
* The Right graph displays the seasonal change of the hiring process. There was a peak(**March**) of the hiring and then it dropped dramatically, and the other small peak occurred right after around **June/July**, which means that different times of a year will influence in the hiring situations

## Part 4: Bar Chart and Targets

```{r barchart.and.targets, echo=F, fig.align='center',fig.height=4.2}
sales<-dat[which(!is.na(dat$`Sales Year to Date`)),]
sales<-sales[which(sales$`Job Title`=="Sales Representative"),]
# Create the data for the chart
sales$H <- sales$`Sales Year to Date`
sales$M <- factor(sales$`First Name`)

sales$flag<-factor(ifelse(sales$`Sales Year to Date`<1600000,"Below New Sales Quota","Above New Sales Quota"))

ggplot(sales,aes(x=M,y=H, fill=flag)) + 
  geom_bar(stat="identity",width = 0.7) +
  labs(title="Sales Representatives & their Sales Year to Date", x="",y="Sales YTD") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(size=8,angle = 30),
        legend.position="top",legend.title=element_blank()) +
  scale_fill_manual(values=c("#00CCCC", "#FF6666")) +
  geom_hline(yintercept = 1600000,col="red",linetype=2) +
  annotate(geom="text", label="New Quota Line", x=1.6, y=1800000, vjust=-0.5, col="red",size=3.4) +
  annotate(geom="text", label="=1600K", x=1.6, y=1600000, vjust=-0.5, col="red",size=3.4)
```

The above shows the bar chart of the Sales Representatives and their SALES YEAR TO DATE as well as the New Sales Quota for them.

## Part 5: Points
The chart shows Salary Rate over time (Day) based on Hire Date and show the points by Gender (Male
or Female) color the Point Pink for Girls and Blue for Boys. 

```{r Points,echo=F, fig.align='center',fig.height=3.5}
time<-dat$`Hire Date`
time<-ymd(time)
dat$timediff<-as.numeric(difftime("2019-10-24",time,units=c("days")))

ggplot(aes(x = timediff, y=Rate, col=Gender),data=dat) + 
  geom_point() +
  labs(title="Salary Rate over time(Day) based on Hire Date", y="Rate", x="Time(Day)") +
  theme(plot.title = element_text(hjust = 0.5), legend.position="right") +
  scale_color_manual(labels = c("Girls", "Boys"), values = c("hotpink1", "steelblue2")) +
  annotate(geom="text", label="High Rate", x=6000, y=115, col="steelblue2",size=6) +
  annotate(geom="text", label="Low Rate", x=6900, y=25, col="steelblue2",size=6)
```

* Generally speaking, the longer an employee was hired, the higher the rate will be
* However, there are some **Low Rate** for the employees who were hired for a long time

## Part 6 Other Insights

In order to investigate the composition of the firm, I mutually divided the employees into 5 departments:**Chief, Research, Finance, Human Resources and Product**.

```{r Others1, echo=F}
job.title <- as.character(unique(dat$`Job Title`))
class <- data.frame(`Job Title` = job.title,
                    "Class" = rep(NA, length(job.title)))
class[c(1:5),2] <- "Chief"
class[c(6:7,11,13,17,19,22,27), 2] <- "Research"
class[c(8:10,16,21,23,28:35,38,39,41,45,47,50:52,54,56:60,62,64,65,67), 2] <- "Product"
class[c(14,15,20,24,25,36,37,40,44,46,49,53,61,63,66), 2] <- "Human\nResources"
class[c(12,18,26,42,43,48,55), 2] <- "Finance"
colnames(class) <- c("Job Title", "Class")
dat.c <- merge(dat, class, by="Job Title")
```

```{r Others2, echo=F, fig.align='center',fig.height=3,fig.width=4.5}
temp<-data.frame(data.frame(table(dat.c$Class)))
packing <- circleProgressiveLayout(temp$Freq, sizetype='area')
data <- cbind(temp, packing)
 
ggplot() + 
  geom_polygon(data = circleLayoutVertices(packing, npoints=50), 
               aes(x, y, group = id, fill=as.factor(id)), colour = "black", alpha = 0.6) +
  geom_text(data = data, aes(x, y, size=100000, label = temp$Var1)) +
  scale_size_continuous(range = c(1,4)) +
  theme_void() + 
  theme(legend.position="none",plot.title = element_text(hjust = 0.5)) +
  coord_equal()
```


The above graph shows that **Product** Department has the largest number of employees, while the **Chief** Office has the smallest number of employees.

```{r Others3, echo=F, fig.align='center',fig.height=4.4,fig.width=7}
temp2<-data.frame(dat.c$Class,dat.c$Gender,dat.c$Rate)
temp3<-ddply(temp2, .(dat.c.Class,dat.c$Gender), summarize, 
             mean = round(mean(dat.c.Rate), 2))
colnames(temp3)<-c("Department","Gender","Average Rate")
ggplot(data=temp3, aes(x=Department, y=`Average Rate`, fill=temp3$Gender)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black") +
    scale_fill_manual(values=c("hotpink1", "steelblue2"),name="Gender") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    labs(x="Department Name",y="Average Rate",title="Average Rate for Female/Male in different departments")
```

This graph shows the average rate for Female and Male from different departments. It shows that:

* The **Chief** Office and **Research** department earn more than the other departments based on the average rate

* **Product** department earns the least in average rate while having the largest number of employees in the firm

* This graph also verified the result I showed before: Women earn more generally and men earn more when he is in a high-level position
